Evaluation of classical correlation functions from 2/3D images on CPU and GPU architectures: introducing CorrelationFunctions.jl

### 2023-05-02 paper-correlations.tex
Do not mention value noise

\section{Efficiency of our algorithms}
\label{sec:efficiency}
To measure efficiecy of our algorithms we use a machine with the following
configuration:
\begin{itemize}
\item \textbf{CPU}: AMD Ryzen 5 1600X
\item \textbf{Memory}: 32~Gb DDR4-2400 memory
\item \textbf{GPU}: Nvidia Tesla V100 with 32~Gb video memory
\end{itemize}

We generate two (resp. three) dimensional square (resp. cubic) binary arrays
with different number of elements. Each element $a_{ij}$ (resp. $a_{ijk}$) of a
two-dimensional (resp. three-dimensional) array is calculated as follows:
\begin{align*}
  a_{ij} & = \left\{
  \begin{array}{ll}
    0 & \quad vn(5i/s, 5j/s, 0, 5, 34365) < 0.5 \\
    1 & \quad \text{otherwise}
  \end{array}
  \right. \\
  a_{ijk} & = \left\{
  \begin{array}{ll}
    0 & \quad vn(5i/s, 5j/s, 5k/s, 5, 34365) < 0.5 \\
    1 & \quad \text{otherwise}
  \end{array}
  \right.
\end{align*}
Here $vn(x,y,z,o,s)$ is a value noise function. We fix a number of
octaves $o = 5$ and a random seed $s = 34365$. The particular implementation of
value noise for our measurements was taken from \code{ValueNoise.jl} package. An
example of a two-dimensional image generated in this way is on
\cref{fig:valuenoise}.

To measure efficiency of \code{Directional} module we generate ten arrays with
a side of a square $s = 1000, 2000, \dots, 10000$ pixels for two-dimensional
case and ten arrays with a side of a cube $s = 50, 100, \dots, 500$ voxels for
three-dimensional case and calculate execution time for each function. All
functions are calculated with default values of optional arguments (that is:
closed walls boundary conditions and computation along axial directions).

With the \code{Map} module we follow the same procedure. Many functions in
\code{Map} can utilize Nvidia's GPUs using the CUDA library. For example,
two-point correlation function uses FFT transform optimized for Nvidia GPUs and
provided in \code{CUDA.FFT} package. We provide execution times for both CPU and
GPU. The choice between CPU or GPU version is made depending on type of input
array. If it's \code{CuArray} rather than generic \code{AbstractArray} then CUDA
implementation is chosen. Please note, that CPU versions are much slower, so we
choose a different sides of squares ($s = 700, 1350, \dots, 7000$) and cubes
($s = 10, 45, \dots, 350$) for evaluation of the performance.

The measurements of execution time is on \cref{fig:timings}. As you can see, GPU
versions (when available) are faster than CPU versions and limited only by
amount of video memory. Functions in \code{Directional} are slower (with
exception of the lineal-path function) but are also useful in medium
reconstruction algorithms which utilize simulated annealing.

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.9\linewidth]{images/timing-image.png}
  \caption[]{An example of 2D image used in timing measurements.}
  \label{fig:valuenoise}
\end{figure}

>>>>

\section{Execution times of our algorithms}
\label{sec:efficiency}
To measure execution times of our algorithms we use a machine with the following
configuration:\textcolor{red}{Must be configuration of Asus notebook}
\begin{itemize}
\item \textbf{CPU}: AMD Ryzen 5 1600X
\item \textbf{Memory}: 32~Gb DDR4-2400 memory
\item \textbf{GPU}: Nvidia Tesla V100 with 32~Gb video memory
\end{itemize}

We generate two (resp. three) dimensional square (resp. cubic) binary arrays
with different sides. An array used for measurements contains two solid disks
(resp. balls) surrounded by void phase, therefore it is excected for $C_2$
function to perform two times slower than $S_2$.

To measure efficiency of \code{Directional} module we generate ten arrays with
a side of a square $s = 1000, 2000, \dots, 10000$ pixels for two-dimensional
case and ten arrays with a side of a cube $s = 50, 100, \dots, 500$ voxels for
three-dimensional case and calculate execution time for each function. All
functions are calculated along axial directions using periodic boundary
conditions.

With the \code{Map} module we follow the same procedure. Many functions in
\code{Map} can utilize Nvidia's GPUs using the CUDA library. For example,
two-point correlation function uses FFT transform optimized for Nvidia GPUs and
provided in \code{CUDA.FFT} package. We provide execution times for both CPU and
GPU. The choice between CPU or GPU version is made depending on type of input
array. If it's \code{CuArray} rather than generic \code{AbstractArray} then CUDA
implementation is chosen. \textcolor{red}{Please note, that CPU versions are
  much slower, so we choose a different sides of squares
  ($s = 700, 1350, \dots, 7000$) and cubes ($s = 10, 45, \dots, 350$)
  for evaluation of the performance.}

The measurements of execution time is on \cref{fig:timings}. As you can see, GPU
versions (when available) are faster than CPU versions and limited only by
amount of video memory. Functions in \code{Directional} are the slowest among
other implementations, but require less memory.

% Replace by two disks?
\begin{figure}[ht]
  \centering
  \includegraphics[width=0.6\linewidth]{images/timing-image.png}
  \caption[]{One of 2D image used in timing measurements.}
  \label{fig:valuenoise}
\end{figure}
